---
title: "esm244_lab10_LB"
author: "Lydia Bleifuss"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      connection = "lter_db") #THIS will automatically connect to your database in all code chunks (including sql) 

library(here)
library(tidyverse)
library(DBI)
library(RSQLite)
```


### A. Connect to the database and tell R we're usign SQLite
```{r}

lter_db <- DBI::dbConnect(drv = RSQLite::SQLite(),
                          dbname = here("data","marine.sqlite"))

```


### B. What tables exist in the database? 

To check qhihc tables exsits (table names) in database, use 'DBI::dbListTables()':
```{r}

DBI::dbListTables(lter_db)

#Should populate [1] "fish_data"     "inverts_data"  "lobsters_data"

```

### C. Which variables are in a database table? 

Use 'DBI::dbListFields()'

```{r}
DBI:: dbListFields(lter_db, "fish_data") 

#[1] "year"        "site"        "common_name" "total_count"

DBI:: dbListFields(lter_db, "inverts_data") 

#[1] "month"       "site"        "common_name" "2016"        "2017"       [6] "2018"

DBI:: dbListFields(lter_db, "lobsters_data") 

#[1] "year"      "month"     "date"      "site"      "transect"  "replicate" [7] "size_mm"

```


#### Read in a database table as an R data frame

Use 'DBI::dbReadTable()' 

```{r}
fish_df <- DBI::dbReadTable(lter_db, "fish_data") 
View(fish_df) #Now can read in dfs from SQLite info and work with it there

```

#### Query a database using SQL 

- SELECT: gets columns
- FROM: what table in a database to get something from
- Use * to indicate "get all" of something (e.g. columns)

###SQL CODE DOWN HERE!!
```{sql connection=lter_db}
SELECT * FROM fish_data
```

To get a subset of the colums, we can just list them after SELECT, separated by a comma: 

#didn't need "connection=" up here because I set it at the top
```{sql}

SELECT year, site FROM fish_data

```

We might want to only query for ROWS (observations) that match a given condition we set. 

- in R: dplyr::filter()
- in SQL: WHERE

```{sql}

SELECT * FROM fish_data WHERE year == 2016

```


Now let's query to keep all columns from fish_data, but only if the common_name variable matches "garibaldi": 

```{sql}

SELECT common_name, site FROM fish_data WHERE common_name == "garibaldi" 

```

In dplyr:
filter for an "and" statement: , & 
SQL: "and" 

Example: Query to only keep observations where the common_name == "garibaldi" AND site == "mohk" 

```{sql}

SELECT * FROM fish_data WHERE common_name == "garibaldi" AND site == "mohk" 

```


To write an OR statement 
- | (vertical line), %in%

Example: Query to keep observations where year is 2017 OR site is "abur" 

```{sql}

SELECT * FROM fish_data WHERE year == 2017 OR site == "abur"

```

#### Store an output from SQL querying 

Add in code chunk header 'output.var = "name_object"'

#in the {}, put output.var = "fish_2017_df" which makes it a dataframe, can now work with this in r codechunk! 
```{sql, output.var = "fish_2017_df"}

SELECT * FROM fish_data WHERE year == 2017

```


#### Query using SQL from an R code chunk 

'DBI::dbSendQuery()'

```{r}
garibaldi <- dbSendQuery(lter_db, "SELECT * FROM fish_data WHERE common_name == 'garibaldi'") #step before making a df

#puttig SQL code into "" so R knows we are using other code 

gar_df <- dbFetch(garibaldi) #This makes the df! 

```


### Querying using dplyr-ish code

- 'dplyr' 


```{r}
fish_tbl <- tbl(lter_db, "fish_data")
```

# Some wrangling: 

```{r}
yearly_fish <- fish_tbl %>% 
  group_by(year) %>% 
  summarize(
    total_fish = sum(total_count)
  )

yearly_fish %>% show_query() #THIS shows you exactly what SQL code you would need to use to do the same thing you just did in dplyr 

yearly_fish %>%  collect() #This gives you the outcome of the SQL code 
```


#### Adding or removing tables from a .sqlite database

USE 'DBI::dbWriteTable' and 'DBI::dbRemoveTable' functions to add or remove tables from/to an existing database

```{r}
dbWriteTable(lter_db, "mtcars_df", mtcars)

# Check to see what tables are in the database now: 

DBI:: dbListTables(lter_db) #now it shows you that mtcars_df in included in the Table 

## dbRemoveTable (now we want to get rid of it again)
DBI::dbRemoveTable(lter_db, "mtcars_df")

DBI::dbListTables(lter_db)
```

